Adaptive Sequential Model Selection

### 2015-05-13 paper.tex
added more discussion of single-step selective p-values

There is some ambiguity involved in deciding how to generalize $z$- and $t$-tests to the selective case. For example, \citet{gsell2013sequential} describe three different null hypotheses that we could consider testing at step $k$ in the case of linear regression under the upper model $M_\infty:\; Y\sim \cN(\mu, \sigma^2 I)$:
\begin{description}
\item[Complete Null:] The previous model, $M_{k-1}$, is correct; i.e., that $\mu = X_{E_{k-1}}\beta$ for some $\beta$.
\item[Incremental Null:] $M_{k-1}$ may not be correct, but $M_k$ is no improvement: i.e., that $X_EX_E\mu = \proj_{X_{E_{k-1}}}\mu$
\end{description}
. \citet{fithian2014optimal} discusses several different tests

@@

The random variable  $p_{k,M}(Y)$ is a valid {\em selective $p$-value} for $M$ if it is stochastically larger than uniform under sampling distributions $F\in M$, given that $M$ is selected. That is,
\[
\P_F\left(p_{k,M}(Y) \leq \alpha \mid M_{k-1}(Y) = M\right) 
\leq \alpha, \quad \forall F\in M, \alpha \in [0,1].
\]
Given selective $p$-values for each fixed candidate pair, we can construct a combined $p$-value for the random null $M_{k-1}(Y)$: 
\[
p_k(y) = p_{k, M_{k-1}(y)}(y),
\]
which is a valid $p$-value on the event $\{F \in M_{k-1}(Y)\}$:
\[
\P_F\left(p_k \leq \alpha \mid M_{k-1}, 
  \;F\in M_{k-1}\right) \leq \alpha, \quad \forall \alpha \in [0,1].
\]
Recall that, as usual $M_{k-1}$ is random and not $F$.

@@

@@

\subsection{Comparison of Selected- and Saturated-Model inference}

@@

As we see below, a similar result holds for the LASSO ever-active path. In fact, it holds for a much larger class of $\ell_1$ regularized exponential family models including the graphical LASSO, group LASSO, and... \WFcomment{more? group LASSO not really $\ell_1$ penalty, but the result is true for group LASSO}

@@

>>>>

There is some ambiguity involved in deciding how to generalize $z$- and $t$-tests to the selective case. For example, \citet{gsell2013sequential} describe three different null hypotheses that we could consider testing at step $k$ in the case of linear regression under the upper model $M_\infty:\; Y\sim \cN(\mu, \sigma^2 I)$:
\begin{description}
\item[Complete Null:] The previous model, $M_{k-1}$, is correct; i.e., that $\mu = X_{E_{k-1}}\beta$ for some $\beta$.
\item[Incremental Null:] $M_{k-1}$ may not be correct, but $M_k$ is no improvement: i.e., that $X_EX_E^\dagger\mu = \proj_{E_{k-1}}\mu$
\end{description}
. \citet{fithian2014optimal} discusses several different tests

@@

The random variable  $p_{k,M}(Y)$ is a valid {\em selective $p$-value} for $M$ at step $k$ if it is stochastically larger than uniform under sampling distributions $F\in M$, given that $M$ is selected. That is,
\[
\P_F\left(p_{k,M}(Y) \leq \alpha \mid M_{k-1}(Y) = M\right) 
\leq \alpha, \quad \forall F\in M, \alpha \in [0,1].
\]
Given selective $p$-values for each fixed candidate pair, we can construct a combined $p$-value for the random null $M_{k-1}(Y)$. Define
\[
p_k(y) = p_{k, M_{k-1}(y)}(y),
\]
which is a valid $p$-value on the event $\{F \in M_{k-1}(Y)\}$:
\[
\P_F\left(p_k \leq \alpha \mid M_{k-1}, 
  \;F\in M_{k-1}\right) \leq \alpha, \quad \forall \alpha \in [0,1].
\]
Recall that $M_{k-1}$ is random and not $F$.

@@

\subsection{Selective $p$-Values in Exponential Families}

@@

In linear regression, there are several tests that we might perform at step $k$ if we are , which generally use

\subsubsection{Saturated-Model Tests}
Most of the work in selective inference for linear regression has focused on inference for least-squares parameters of the model $M_\infty:\; Y \sim \cN(\mu, \sigma^2I)$, with $\sigma^2$ assumed known. The saturated-model test for significance of variable $j_k$ after adjusting for the variables in $E_{k-1}$ tests whether the coefficient of $X_{j_k}$ is nonzero in the best linear model $X_{E_k}^\dagger Y$.

Let $\proj_{E_{k-1}}\mu$ denote the projection of $\mu$ onto the column space of $X_{E_{k-1}}$, and let $\eta = \proj_{E_{k-1}}^\perp X_{j_k}$. Saturated-model $p$-values test the null hypothesis $H_0:\;\eta'\mu = 0$. Because $M_\infty$ has $n$ parameters, we must  condition on $\proj_\eta^\perp Y$ to eliminate the effect of the nuisance parameters $\proj_\eta^\perp \mu$ from the problem. Thus, the test is based on
\begin{equation}\label{eq:satModel}
\L\left( \eta'Y \mid E_{k-1}, j_k, \proj_\eta^\perp Y \right).
\end{equation}
To carry out exact tests of this hypothesis, we must assume that $\sigma^2$ is known, or that an independent estimate is available.

\WFcomment{References.} \citet{fithian2014optimal} call $M_\infty$ the {\em saturated model} because there is a parameter for every data point. They contrast tests based on $M_\infty$ with {\em selected-model} tests, which do not treat $\proj_\eta^\perp \mu$ as a nuisance parameter.

\subsubsection{Selected-Model Tests}
Selected-model inference~\citep{fithian2014optimal} presents a more powerful alternative to testing for inclusion of variable $j_k$. Selected-model tests are so called because they are based on the statistical model chosen by our selection procedure. In this case, the null model selected for testing at step $k$ is simply the model $M_{k-1}:\; Y \sim \cN(X_{E_{k-1}}\beta, \sigma^2 I)$.

There is a UMPU test of the null $M_{k-1}$ against the alternative $M_{k}\setminus M_{k-1}$, which has one more degree of freedom than $M_{k-1}$. One option for selected-model inference is to condition on the identity of $j_k$, the next variable added, and perform the UMPU selective test of $M_{k-1}$ against $M_{k}\setminus M_{k-1}$. In that case, the test is based on
\begin{equation}\label{eq:selModel_cond}
\L\left(X_{j_k}'Y \mid E_{k-1}, \; j_k, \; X_{E_{k-1}}'Y\right).
\end{equation}
Note the contrast between~\eqref{eq:selModel_cond} and~\eqref{eq:satModel}. In~\eqref{eq:selModel_cond}, we only need to condition on an $|E_{k-1}|$-dimensional projection of $Y$, whereas in~\eqref{eq:satModel} we condition on $n-1$-dimensional projection of $Y$. Conditioning on more information tends to sap the power of a test, and we will see in Sections~\ref{sec:bivariate} and~\ref{sec:sparseReg}, the saturated-model tests often pay a heavy price for this extra conditioning.

Although the tests in~\eqref{eq:selModel_cond} and~\eqref{eq:satModel} may appear to be based on different test statistics, they are actually not: because $\eta'Y = X_{j_k}'(Y - \proj_{E_{k-1}}Y)$, one test statistic is a deterministic affine function of the other after we condition on $X_{E_{k-1}}'Y$.

It is not really necessary to condition on $j_k$: instead we could use
\begin{equation}\label{eq:selModel_marg}
\L\left(X_{j_k}'Y \mid E_{k-1}, \; X_{E_{k-1}}'Y\right),
\end{equation}
or even replace $X_{j_k}'Y$ with any other test statistic. Under the null hypothesis, the distribution of $Y$ is completely known once we condition on the value of $X_{E_{k-1}}'Y$, the complete sufficient statistic of $M_{k-1}$.

We will see in Section~\ref{sec:pValsIndep} that conditioning on $j_k$ as in~\eqref{eq:selModel_cond} always results in independent sequential $p$-values, whereas tests of the form~\eqref{eq:selModel_marg} yield independent $p$-values only under certain conditions. Section~\ref{sec:pValsIndep} explores the conditions required to obtain independent sequential $p$-values. Carrying out a finer comparison of the two selected-model tests is an interesting direction for future work.

Selected-model inference is more computationally involved than saturated-model inference. Because the saturated-model test conditions on $n-1$ dimensions, the resulting distribution in~\eqref{eq:satModel} is nothing more than a truncated univariate Gaussian random variable.

However, there are several major benefits to selected-model inference that often justify its use. First, selected-model inference allows us to drop the assumption that $\sigma^2$ is known. Second, $p$-values for selected-model tests are independent under general conditions, allowing us to plug the $p$-values into powerful sequential stopping rules such as those developed by~\citet{gsell2013sequential}. Most importantly, as we will see, selected-model tests can be dramatically more powerful than saturated-model tests. See \citet[][Section 5]{fithian2014optimal} for a detailed discussion of the difference between saturated-model and selected-model tests.

\subsection{Comparison of Selected- and Saturated-Model inference}\label{sec:bivariate}

@@

As we see below, a similar result holds for the LASSO ever-active path. In fact, it holds for a much larger class of $\ell_1$-regularized exponential family models including the graphical LASSO.

@@

\WFcomment{Give more general characterization of when we can do it, including group lasso? QP? Generalized Lasso?}
